---
title: 'NLP workshop - Exploring Presidential Debate on twitter'
author: "Daniel S. Hain (dsh@business.aau.dk)"
date: "Updated `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    code_folding: show
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
    theme: flatly
---

```{r setup, include=FALSE}
### Generic preamble
rm(list=ls())
Sys.setenv(LANG = "en") # For english language
options(scipen = 5) # To deactivate annoying scientific number notation

### Knitr options
library(knitr) # For display of the markdown
knitr::opts_chunk$set(warning=FALSE,
                     message=FALSE,
                     comment=FALSE, 
                     fig.align="center"
                     )
```

```{r}
### Load standardpackages
library(tidyverse) # Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) # For extra-piping operators (eg. %<>%)
```

```{r}
library(tidytext)
```


# Download the data

```{r}
# download and open some Trump tweets from trump_tweet_data_archive
library(jsonlite)
tmp <- tempfile()
download.file("https://github.com/SDS-AAU/SDS-master/raw/master/M2/data/pol_tweets.gz", tmp)

tweets_raw <- stream_in(gzfile(tmp, "pol_tweets"))
```

```{r}
tweets_raw %>% glimpse()
```

```{r}
tweets <- tibble(ID = colnames(tweets_raw[[1]]), 
                 text = tweets_raw[[1]] %>% as.character(), 
                 labels = tweets_raw[[2]] %>% as.logical())
#rm(tweets_raw)
```

```{r}
tweets %>% glimpse()
```

```{r}
tweets %<>%
  filter(!(text %>% str_detect('^RT'))) # Filter retweets
```

```{r}
tweets %>% head()
```

# Tidying

```{r}
tweets_tidy <- tweets %>%
  unnest_tokens(word, text, token = "tweets") 
```

```{r}
tweets_tidy %>% head(50)
```


```{r}
tweets_tidy %>% count(word, sort = TRUE)
```


# Preprocessing

```{r}
# preprocessing
tweets_tidy %<>%
  filter(!(word %>% str_detect('@'))) %>% # remove hashtags and mentions
  filter(!(word %>% str_detect('^amp|^http|^t\\.co'))) %>% # Twitter specific stuff
#  mutate(word = word %>% str_remove_all('[^[:alnum:]]')) %>% ## remove all special characters
  filter(str_length(word) > 2 ) %>% # Remove words with less than  3 characters
  group_by(word) %>%
  filter(n() > 100) %>% # remove words occuring less than 100 times
  ungroup() %>%
  anti_join(stop_words, by = 'word') # remove stopwords
```

# TFIDF

TFIDF weighting

```{r}
# top words
tweets_tidy %>%
  count(word, sort = TRUE) %>%
  head(20)
```

```{r}
# TFIDF weights
tweets_tidy %<>%
  add_count(ID, word) %>%
  bind_tf_idf(term = word,
              document = ID,
              n = n)
```


```{r}
# TFIDF topwords
tweets_tidy %>%
  count(word, wt = tf_idf, sort = TRUE) %>%
  head(20)
```

# Inspecting

## Words by party affiliation

```{r}
labels_words <- tweets_tidy %>%
  group_by(labels) %>%
  count(word, wt = tf_idf, sort = TRUE, name = "tf_idf") %>%
  slice(1:100) %>%
  ungroup() 
```

```{r, fig.width=10}
labels_words %>%
  mutate(word = reorder_within(word, by = tf_idf, within = labels)) %>%
  ggplot(aes(x = word, y = tf_idf, fill = labels)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~labels, ncol = 2, scales = "free") +
  coord_flip() +
  scale_x_reordered()
```


```{r}
words_classifier <- labels_words %>%
  arrange(desc(tf_idf)) %>%
  distinct(word, .keep_all = TRUE) %>%
  select(-tf_idf)
```

```{r}
tweet_null_model <- tweets_tidy %>%
  inner_join(labels_words, by = 'word')
```

```{r}
null_res <- tweet_null_model %>%
  group_by(ID) %>%
  summarise(truth = mean(labels.x, na.rm = TRUE) %>% round(0),
         pred = mean(labels.y, na.rm = TRUE) %>% round(0))
```

```{r}
table(null_res$truth, null_res$pred)
```
```{r}
tweets %>% count(labels)
```



# Predictive model

```{r}
library(tidymodels)
```


```{r}
# Notice, we use the initial untokenized tweets
data <- tweets %>%
  select(labels, text) %>%
  rename(y = labels) %>%
  mutate(y = y  %>% as.factor()) 
```


## Training & Test split

```{r}
data_split <- initial_split(data, prop = 0.75, strata = y)

data_train <- data_split  %>%  training()
data_test <- data_split %>% testing()
```

## Preprocessing pipeline

```{r}
library(textrecipes) # Adittional recipes for working with text data
```

```{r}
# This recipe pretty much reconstructs all preprocessing we did so far
data_recipe <- data_train %>%
  recipe(y ~.) %>%
  themis::step_downsample(y) %>% # For downsampling class imbalances (optimal)
  step_filter(!(text %>% str_detect('^RT'))) %>% # Upfront filtering retweets
  step_filter(text != "") %>%
  step_tokenize(text, token = "tweets") %>% # tokenize
  step_tokenfilter(text, min_times = 100) %>%  # Filter out rare words
  step_stopwords(text, keep = FALSE) %>% # Filter stopwords
  step_tfidf(text) %>% # TFIDF weighting
#  step_pca(all_predictors()) %>% # Dimensionality reduction via PCA (optional)
  prep() # NOTE: Only prep the recipe when not using in a workflow
```

```{r}
data_recipe
```

Since we will not do hyperparameter tuning, we directly bake/juice the recipe

```{r}
data_train_prep <- data_recipe %>% juice()
data_test_prep <- data_recipe %>% bake(data_test)
```


## Defining the models

```{r}
model_en <- logistic_reg(mode = 'classification', 
                         mixture = 0.5, 
                         penalty = 0.5) %>%
  set_engine('glm', family = binomial) 
```


## Define the workflow

We will skip the workflow step this time, since we do not evaluate different models against each others.

## fit the model

```{r}
fit_en <- model_en %>% fit(formula = y ~., data = data_train_prep)
```


```{r}
pred_collected <- tibble(
  truth = data_train_prep %>% pull(y),
  pred = fit_en %>% predict(new_data = data_train_prep) %>% pull(.pred_class),
  pred_prob = fit_en %>% predict(new_data = data_train_prep, type = "prob") %>% pull(.pred_TRUE),
  ) 
```

```{r}
pred_collected %>% conf_mat(truth, pred)
```

```{r}
pred_collected %>% conf_mat(truth, pred) %>% summary()
```

Well... soso

# Using the model for new prediction

## Simple test

```{r}
# How would the model predict given some tweet text
pred_own = tibble(text = 'USA USA WE NEED A WALL TO MAKE AMERICA GREAT AGAIN AND KEEP THE MEXICANS AND ALL REALLY BAD COUNTRIES OUT! AMNERICA FIRST')
```


```{r}
fit_en %>% predict(new_data = data_recipe %>% bake(pred_own))
```


## New data

* We could also use the model to predict on new data, such as the just scraped discussion on the presidential debate.

```{r}
# download and open some Trump tweets from trump_tweet_data_archive
library(jsonlite)
tmp <- tempfile()
download.file("https://github.com/SDS-AAU/SDS-master/raw/master/M2/data/pres_debate_2020.gz", tmp)

tweets_raw <- stream_in(gzfile(tmp, "pres_debate_2020"))
```

```{r}
tweets_raw %>% glimpse()
```

```{r}
tweets_new <- tibble(ID = tweets_raw[[1]]$id, 
                     text = tweets_raw[[1]]$tweet)
rm(tweets_raw)
```

```{r}
pred_new <- fit_en %>% predict(new_data = data_recipe %>% bake(tweets_new))
```

To be continued by you :)


# Topic models (LDA)

```{r}
# for LDA analysis
library(topicmodels)
```

### Preparing the Data

```{r}
# LDA via the topicmodel package requires a document-term-matrix (dtm)
tweets_dtm <- tweets_tidy %>%
  cast_dtm(document = ID, term = word, value = n)
```

Lets take a look:

```{r}
tweets_dtm
```

* We see again hat the matrix is still rather sparse, which is an artefact of text data generally, but even more so when using twitter data. 
* Lets try to see if we could reduce that somewhat by deleting less often used terms.

```{r}
library(tm)
tweets_dtm %>% removeSparseTerms(sparse = .99)
```

```{r}
tweets_dtm %>% removeSparseTerms(sparse = .999)
```

```{r}
tweets_dtm %>% removeSparseTerms(sparse = .9999)
```

* Ok, we might have to accept a high level of sparsity in order to still have a meaningful number of unique words.
* Now we can perform a LDA, using the more accurate Gibbs sampling as `method`.

```{r}
tweets_lda <- tweets_dtm %>% 
  LDA(k = 6, method = "Gibbs",
      control = list(seed = 1337))
```

### $\beta$: Word-Topic Association

* $\beta$ is an output of the LDA model, indicating the propability that a word occurs in a certain topic.
* Therefore, loking at the top probability words of a topic often gives us a good intuition regarding its properties.

```{r}
# LDA output is defined for tidy(), so we can easily extract it
lda_beta <- tweets_lda %>% 
  tidy(matrix = "beta") %>%
  group_by(topic) %>%
  arrange(topic, desc(beta)) %>%
  slice(1:10) %>%
  ungroup() 
```

```{r}
lda_beta %>% head()
```

```{r}
# Notice the "reorder_within()"
lda_beta %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  group_by(topic, term) %>%    
  arrange(desc(beta)) %>%  
  ungroup() %>%
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Top 10 terms in each LDA topic",
       x = NULL, y = expression(beta)) +
  facet_wrap(~ topic, ncol = 2, scales = "free")
```

### $\gamma$: Document-Topic Association

* In LDA, documents are represented as a mix of topics. This association of a document to a topic is captured by $\gamma$

```{r}
lda_gamma <- tweets_lda %>% 
  tidy(matrix = "gamma")
```

```{r}
lda_gamma %>% head()
```

```{r}
lda_gamma %>%
  ggplot(aes(gamma)) +
  geom_histogram() +
  scale_y_log10() +
  labs(title = "Distribution of probabilities for all topics",
       y = "Number of documents", x = expression(gamma))
```

```{r}
lda_gamma %>% 
  group_by(topic) %>%
  arrange(desc(gamma)) %>%
  slice(1:3) %>%
  ungroup() %>%
  left_join(tweets %>% select(ID, text), by = c('document' = 'ID')) %>%
  select(topic, text)
```


```{r}
lda_gamma %>%
  ggplot(aes(gamma, fill = as.factor(topic))) +
  geom_histogram(show.legend = FALSE) +
  facet_wrap(~ topic, ncol = 2) +
  scale_y_log10() +
  labs(title = "Distribution of probability for each topic",
       y = "Number of documents", x = expression(gamma))
```

```{r}
top_topics <- tweets_lda %>% 
  tidy(matrix = "gamma")  %>%
  group_by(document) %>%
  top_n(1, wt = gamma) %>%
  ungroup()
```

```{r}
top_topics %>%
  count(topic)
```

```{r}
x <- lda_gamma %>%
  left_join(tweets %>% select(ID, labels), by = c('document' = 'ID'))
```

```{r}
x %>%
  group_by(labels, topic) %>%
  summarise(n = sum(gamma) ) %>%
  group_by(labels) %>%
  mutate(n = n / sum(n)) %>%
  ungroup()
```

# Endnotes

### Packages & Ecosystem

* [`tidytext`](https://github.com/juliasilge/tidytext)
* [`textrecipes`](https://textrecipes.tidymodels.org/)
* [`topicmodels`](https://cran.r-project.org/web/packages/topicmodels/vignettes/topicmodels.pdf)

Further NLP packages ecosystem

* `tm` [here](https://cran.r-project.org/web/packages/tm/)
* `quanteda` [here](https://quanteda.io/), and many many great tutorials [here](https://tutorials.quanteda.io/)


### References 

* Julia Silge and David Robinson (2020). Text Mining with R: A Tidy Approach, O’Reilly. Online available [here](https://www.tidytextmining.com/)
   * [Chapter 6](https://www.tidytextmining.com/topicmodeling.html): Introduction topic models
* Emil Hvidfeldt and Julia Silge (2020). Supervised Machine Learning for Text Analysis in R, online available [here](https://smltar.com/)
   * [Chapter 7](https://smltar.com/mlclassification.html): Classification

### Further sources

Datacamp

*  [Topic Modeling in R](https://learn.datacamp.com/courses/topic-modeling-in-r) 

Other online

* [Julia Silge's Blog](https://juliasilge.com/): Full of great examples of predictive modeling, NLP, and the combination fo both, using tidy ecosystems

### Session Info

```{r}
sessionInfo()
```

