---
title: "Assignment: Data Manipulation, EDA and Viz - Nomad Dataset (Example Solution)"
author: "Daniel S. Hain (dsh@business.aau.dk)"
date: "Updated `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    code_folding: show
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
    theme: flatly
---

```{r setup, include=FALSE}
# Knitr options
### Generic preamble
Sys.setenv(LANG = "en") # For english language
options(scipen = 5) # To deactivate annoying scientific number notation

# rm(list=ls()); graphics.off() # get rid of everything in the workspace
if (!require("knitr")) install.packages("knitr"); library(knitr) # For display of the markdown

### Knitr options
knitr::opts_chunk$set(warning=FALSE,
                     message=FALSE,
                     fig.align="center"
                     )
```

# Preamble

## Standard packages

```{r}
### Load packages
library(tidyverse) # Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) # For extra-piping operators (eg. %<>%)
```

## Load data

Trips

```{r}
trips <- read_csv('https://sds-aau.github.io/SDS-master/M1/data/trips.csv')
```

```{r}
trips %>% glimpse()
```

People

```{r}
people <- read_csv('https://sds-aau.github.io/SDS-master/M1/data/people.csv')
```

```{r}
people %>% glimpse()
```

Countries

```{r}
countries <- read_csv( 'https://sds-aau.github.io/SDS-master/M1/data/countrylist.csv')
```

```{r}
countries %>% glimpse()
```

# 1: Preprocessing

## a. Trips: transform dates into timestamps

```{r}
trips %>% select(date_start, date_end) %>%
  glimpse()
```

`readr` is smart, so if you loaded the data with `read_csv`, then this is already taken care of. Otherwise:

```{r}
# To demonstrate, I transform it back to a string.
trips %<>% mutate(date_start = date_start %>% as.character(),
                  date_end = date_end %>% as.character())
```

```{r}
trips %>% select(date_start, date_end) %>%
  glimpse()
```

In case it is a string but well formated, we can use the `lubridate` packages.

```{r}
library(lubridate) # This is tidyverse's datetime package

trips %<>% mutate(date_start = date_start %>% as_date(),
                  date_end = date_end %>% as_date())
```

```{r}
trips %>% select(date_start, date_end) %>%
  glimpse()
```

## b. Calculate trip duration in days

```{r}
trips %<>% mutate(trip_duration = date_end - date_start)
```

```{r}
# Test if it works
trips %>% 
  select(trip_duration, date_start, date_end) %>%
  head()
```

Seems to work fine :)

## c. Filter extreme (fake?) observations for durations as well as dates - start and end

Lets inspect:

```{r}
trips %>% 
  select(trip_duration, date_start, date_end) %>%
  summary()
```

We clearly see that some observations areunrealistic (trip in Jesus's times or in the future etc.). Lets look at the distribution

```{r}
trips %>%
  ggplot(aes(x = date_start)) +
  geom_histogram()
```

```{r}
trips %>%
  ggplot(aes(x = date_end)) +
  geom_histogram()
```

There are many ways to deal with outliers. To make it simple:

1: We could filter by some minimum / maximum date set manually
2: We could just delete extreme values using `percentage_rank` (deleting the 1 percent of obs with highest/lowest values). We will demonstrate this here:

```{r}
trips %<>%
  mutate(date_start_pct = date_start %>% as.numeric() %>% percent_rank(),
         date_end_pct = date_end %>% as.numeric() %>% percent_rank()) %>%
  filter(date_start_pct >= 0.01 & date_start_pct <= 0.99) %>%
  filter(date_end_pct >= 0.01 & date_end_pct <= 0.99) 
```

Lets check how it looks now:

```{r}
trips %>% 
  select(trip_duration, date_start, date_end) %>%
  summary()
```

We clearly see that some observations areunrealistic (trip in Jesus's times or in the future etc.). Lets look at the distribution

```{r}
trips %>%
  ggplot(aes(x = date_start)) +
  geom_histogram()
```
```{r}
trips %>%
  ggplot(aes(x = date_end)) +
  geom_histogram()
```

Way more realistic, right?

## d. Join the countrylist data to the trips data-frame using the countrycode as a key

A simple left join.Be only aware of the different variable names

```{r}
trips %<>%
  left_join(countries, by = c("country_code" = "alpha_2"))
```

```{r}
trips %>% head()
```

New variables are in, seems to work. Lets check if there are some trips that could not match:

```{r}
trips %>% filter(is.na(region) | is.na(sub_region))
```

```{r}
trips %>% filter(is.na(region) | is.na(sub_region)) %>%
  count(country_code, sort = TRUE)
```

Ok, we see some country codes did not match. We dont bother for most small numbers, but one thing we might take a look at: UK did not match, since it is coded GB in the countries dataframe (Just inspect it). Lets delete the newly matched variables and start over again.

```{r}
trips %<>% 
  select(-region, -sub_region)
```

Lets replace UK with GB

```{r}
trips %<>% 
  mutate(country_code = country_code %>% str_replace(pattern = 'UK', replacement = 'GB'))
```

```{r}
trips %>%
  filter(country_code == 'UK' |country_code == 'GB') %>%
  count(country_code)
```

Ok, no more UK present... lets join again.

```{r}
trips %<>%
  left_join(countries, by = c("country_code" = "alpha_2"))
```

```{r}
trips %>% filter(is.na(region) | is.na(sub_region)) %>%
  count(country_code, sort = TRUE)
```

Ok, the rest seems negligible... we just delete these observations...

```{r}
trips %<>% drop_na(region)
```

# 2: People

## a: How many people have a least a “High School” diploma?

Lets see what educations we have in the data

```{r}
people %>% count(education_raw)
```

Ok, that seems easy. Since all educations include Highschool (or higher), we can just assume that everybody that has all people with a non-missing education field have at least a highschool degree.

```{r}
people %>% pull(education_raw) %>% sum(!is.na())
```

