---
title: "Workshop: Exploring the InsideAirBnB dataset - EDA"
author: "Daniel S. Hain (dsh@business.aau.dk)"
date: "Updated `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    code_folding: show
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
    theme: flatly
---

```{r setup, include=FALSE}
# Knitr options
### Generic preamble
Sys.setenv(LANG = "en") # For english language
options(scipen = 5) # To deactivate annoying scientific number notation

# rm(list=ls()); graphics.off() # get rid of everything in the workspace
if (!require("knitr")) install.packages("knitr"); library(knitr) # For display of the markdown

### Knitr options
knitr::opts_chunk$set(warning=FALSE,
                     message=FALSE,
                     fig.align="center"
                     )
```

## Preamble

```{r}
# Clear workspace
rm(list=ls()); graphics.off() 
```

```{r}
### Load packages
library(tidyverse) # Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) # For extra-piping operators (eg. %<>%)
library(skimr) # For nice data summaries
```


# The InsideAirBnB data

## Instroduction


* The data is sourced from the [**Inside Airbnb**](http://insideairbnb.com/get-the-data.html) which hosts publicly available data from the Airbnb site.
* Interactive visualizations are provided [here](http://insideairbnb.com/copenhagen/?neighbourhood=&filterEntireHomes=false&filterHighlyAvailable=false&filterRecentReviews=false&filterMultiListings=false)

The dataset comprises of three main tables:

* `listings` - Detailed listings data showing 96 atttributes for each of the listings. Some of the attributes which are intuitivly interesting are: `price` (continuous), `longitude` (continuous), `latitude` (continuous), `listing_type` (categorical), `is_superhost` (categorical), `neighbourhood` (categorical), `ratings` (continuous) among others.
* `reviews` - Detailed reviews given by the guests with 6 attributes. Key attributes include `date` (datetime), `listing_id` (discrete), `reviewer_id` (discrete) and `comment` (textual).
* `calendar` - Provides details about booking for the next year by listing. Four attributes in total including `listing_id` (discrete), `date` (datetime), `available` (categorical) and `price` (continuous).

## Load data

```{r}
listings <- read_csv('http://data.insideairbnb.com/denmark/hovedstaden/copenhagen/2020-06-26/data/listings.csv.gz')
listings %>% head()
```

```{r}
calendar <- read_csv('http://data.insideairbnb.com/denmark/hovedstaden/copenhagen/2020-06-26/data/calendar.csv.gz')
calendar %>% head()
```

```{r}
reviews <- read_csv('http://data.insideairbnb.com/denmark/hovedstaden/copenhagen/2020-06-26/data/reviews.csv.gz')
reviews %>% head()
```

```{r}
# The geodat of the hoods comes as a geojson, so we need the right package to load it
library(geojsonio)
neighbourhoods_geojson <- geojson_read( 'http://data.insideairbnb.com/denmark/hovedstaden/copenhagen/2020-06-26/visualisations/neighbourhoods.geojson',  what = "sp")
```


```{r}
listings %>% skim()
```

```{r}
listings %>% select(price) %>% skim()
```

```{r}
listings %>% select(price) %>% head()
```































## Some preprocessing

```{r}
listings %<>% mutate(price = price %>% parse_number()) 
```

# Ad-hoc investigation

## Problem 1: Professional hosts & their characteristics

```{r}
listings %>%
  count(host_id, sort = TRUE)
```

Where are they?

```{r}
listings %>%
  filter(host_id == 187610263) %>%
  count(neighbourhood_cleansed, sort = TRUE)
```
Dummy for professional host
```{r}
listings %<>%
  group_by(host_id) %>%
  mutate(host_professional = n() >= 5) %>%
  ungroup()
```

```{r}
listings %>%
  group_by(host_professional) %>%
  summarise(review = review_scores_rating %>% mean(na.rm = TRUE),
            price = price %>% mean(na.rm = TRUE))
```
-> Profressional hosts charge more...

```{r}
listings %>%
  group_by(neighbourhood_cleansed, host_professional) %>%
  summarise(review = review_scores_rating %>% mean(na.rm = TRUE)) %>%
  pivot_wider(names_from = host_professional, values_from = review)
```

-> This is true everywhere, yet in some hoods mnore tghan in others

## Problem 2: Length of description -> Satisfaction

```{r}
listings %<>%
  mutate(desc_lenght = description %>% str_count('\\w+')) %>%
  mutate(desc_long =  percent_rank(desc_lenght) > 0.9 )
```

```{r}
listings %>%
  group_by(desc_long) %>%
  summarise(review = review_scores_rating %>% mean(na.rm =TRUE))
```
-> No overall effect

## P3: Best party place

```{r}
listings %<>% 
  mutate(party_place = accommodates >= 10) 
```

```{r}
listings %>% 
  filter(party_place == TRUE) %>%
  group_by(neighbourhood_cleansed) %>%
  summarize(n = n(),
         review = review_scores_rating %>% mean(na.rm = TRUE),
         price = price %>% mean(na.rm = TRUE),
         price_pp = (price / accommodates) %>% mean(na.rm = TRUE)) %>%
  arrange(desc(n))
```

If you are on a tight budget, best go to Amager-Vest.

# Geoplotting

## Interactive map

```{r}
library(leaflet)
```

```{r}
listings %>% leaflet() %>%
  addTiles() %>%
  addMarkers(~longitude, ~latitude,
             labelOptions = labelOptions(noHide = F),
             clusterOptions = markerClusterOptions(),
             popup = paste0("<b> Name: </b>", listings$name, 
                            "<br/><b> Host Name: </b>", listings$host_name, 
                            "<br> <b> Price: </b>", listings$price, 
                            "<br/><b> Room Type: </b>", listings$room_type, 
                            "<br/><b> Property Type: </b>", listings$property_type
                 )) %>% 
#  setView(-74.00, 40.71, zoom = 12) %>%
  addProviderTiles("CartoDB.Positron")
```

## Choropleth Plots

```{r}
# Using broom to tidy the geojson
library(broom)
neighbourhoods_tidy <-  neighbourhoods_geojson %>%
  tidy(region = "neighbourhood")
```

```{r}
neighbourhoods_tidy %>% glimpse()
```

```{r}
neighbourhoods_tidy %>%
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon() +
  theme_void() +
  coord_map()
```
```{r}
neighborhood_agg <- listings %>%
  group_by(neighbourhood_cleansed) %>%
  summarise(n = n(),
            price_mean = price %>% mean(na.rm = TRUE),
            review_mean = review_scores_rating %>% mean(na.rm = TRUE))
  
```

```{r}
neighbourhoods_tidy %<>%
  left_join(neighborhood_agg, by = c('id' = 'neighbourhood_cleansed'))
```

Number of places

```{r}
neighbourhoods_tidy %>%
  ggplot(aes(x = long, y = lat, group = group, fill = n)) +
  geom_polygon() +
  theme_void() +
  coord_map()
```
Prices

```{r}
neighbourhoods_tidy %>%
  ggplot(aes(x = long, y = lat, group = group, fill = price_mean)) +
  geom_polygon() +
  theme_void() +
  coord_map()
```
Review scores

```{r}
neighbourhoods_tidy %>%
  ggplot(aes(x = long, y = lat, group = group, fill = review_mean)) +
  geom_polygon() +
  theme_void() +
  coord_map()
```

